\subsection{Improved metrics to quantify differential expression}

Analysis of expression differences between individual groups is the basis of most profiling efforts. Variance-based metrics, such as Analysis of Variance (ANOVA) F-Value, or coefficient of variation (CV) are commonly used for this purpose. These metrics are jointly affected by the information content of the differential expression (pattern) and the robustness of the differences (effect size) and so cannot readily separate these two parameters. As a complement to traditional metrics and to begin mining our extensive and complex dataset for novel insights, we developed two easily calculated metrics that better separate the information content and the robustness of expression differences. 

First, in order to extract the transcriptional signals related to cell type identity, we quantified each gene's ability to differentiate each pair of profiled cell types. Based on expression levels and variability (Figure 3A; Methods) we compiled a Differentiation Matrix (DM) with elements equal to one or zero depending on whether or not the gene is differentially expressed between each pair of profiles. The Differentiation Index (DI) is simply the fraction of pairs distinguished, excluding self-comparisons; and ranges from 0 to 1. The maximum observed value of 0.65 indicates that the gene distinguishes 65\% of the pairs, while a value of 0 indicates that the gene distinguishes none (i.e., expressed at similar levels in all cell types).

The ability to detect transcriptional differences between cell types depends on both magnitude of difference and associated noise. To quantify this in our second metric, we defined the Signal Contrast (SC), which closely reflects Signal-to-Noise-Ratio (SNR). Since the signals we are interested in are the gene expression differences distinguishing cell types, we used a noise estimate derived from all undistinguished pairs from the same gene. SC, which indicates how robustly pairs are distinguished, is the ratio of the average effect size for distinguished and undistinguished pairs. High SC genes robustly distinguish cell populations and are therefore suitable as "marker genes". 

Our metrics outperform existing metrics such as ANOVA, CV, and Fano factor in distinguishing the information content and robustness of differential expression. To illustrate the properties of DI and SC relative to existing metrics, we calculated these metrics against various simulated expression patterns with added noise (Figure 3 Supplement 1A). The results (Figure 3 Supplement 1A, lower part) demonstrate that DI (blue) is highly correlated with mutual information (MI; green), yet much easier to calculate. This makes intuitive sense, since the division of cell types into those that can and cannot be distinguished (DM; Figure 3A) corresponds to a unit of information about cell types provided by a gene expression pattern (for more details of the relationship between DI and MI, see Figure 3 Supplement 1C and 2). 
%As shown below (Methods), in the simplified case of binary expression levels and ignoring noise, DI is an order-preserved transformation of MI, although MI is usually significantly more difficult to calculate precisely. More generally, DI is highly correlated with MI and can be regarded as an approximation to MI (For another metric related to DI, see Simpson's index of diversity \cite{SIMPSON_1949}). 
The simulations also show that DI is fairly independent from SNR. For example, both high and low SNR binary patterns yield similar DIs. In contrast, SC (orange) is independent from MI, but is highly correlated to SNR. Thus, DI provides an estimate of the information content of expression patterns across cell types, whereas SC provides an estimate of SNR.

Unlike DI and SC, traditional variance-based methods like ANOVA F-values and CV are either affected by both MI and SNR (ANOVA) or by neither (CV). These differences between metrics are summarized in Figure 3 Supplement 1B. The fact that ANOVA does not distinguish between information content and SNR is also apparent in the data. As shown in Figure 3B, high-ANOVA genes include both high DI and high SC genes. Therefore, SC and DI are useful because they provide independent measures of the robustness and magnitude of differential expression between cell types.

To determine the types of genes most differentially expressed (highest DI) and most robustly different (highest SC) between cell types, we used the PANTHER \citep{Thomas_2003} gene families (Figure 3C). As expected, high DI genes are enriched for neuronal effector genes including receptors, ion channels and cell adhesion molecules (Figure 3C top). The least noisy expression differences (highest SC) were those of homeobox transcription factors (TFs) and the more inclusive categories (TFs, DNA binding proteins) that encompass them (Figure 3C bottom). Hence DI and SC respectively emphasize the information content of genes  mediating the distinctive neuronal phenotypes that distinguish cell types, and the robust, low-noise expression of genes involved in shaping these cell types unique transcriptional programs.

Genes may also contribute to cell type differences through differential splicing. We analyzed splicing events by computing the relative likelihood (branch probabilities) of each donor site in a transcript being spliced to multiple acceptor sites, and of each acceptor site being spliced to multiple donors (Figure 3 D). Interestingly, when these branch probabilities are computer separately for each cell type, they are highly bimodal, reflecting virtually all-or-none splicing at each alternatively spliced site. This pattern has previously been observed for individual cells in some systems\cite{Shalek_2013}. The present observations suggest that these splicing decisions are made at the level of  cell types, rather than independently for individual cells of the same type. We applied a variant of the DM/DI method to alternative splicing (Figure 3 D-F; for details see Methods) and found that voltage-gated calcium and sodium channels are highly alternatively spliced, consistent with previously known results \citep[e.g.][]{Lipscombe_2013}. We also found that G-protein modulators, especially guanyl-nucleotide exchange factors (GEFs), are highly alternatively spliced. Hence, differential splicing of multi-exon genes also contributes to transcriptome diversity across neuronal cell types.






















