\section{Supplementary Materials}

\subsection{Supplementary Files} 

\textbf{Supplementary File 1} Table listing information for mouse lines. Information (columns) includes regions profiled, source of the mouse line, repository ID and URL, whether atlas is in Janelia viewer, URL for other atlases, and relevant references.

\textbf{Supplementary File 2} Table for sample information. Included fields are, \\
1. sample\_id: Sample ID; \\
2. sample\_name: Sample Name; \\ 
3. group: Sample Group ID; \\
4. group\_label: Label for Group; \\
5. sample\_label: Label for Sample;  \\
6. seqlane: Sequencing Lane ID; \\
7. mouseline: Mouse Line ID; \\
8. sample\_code: Type of sample, cs.n: cell-type-specific neuronal sample; cs.o: cell-type-specific non-neuronal sample; ti.b: tissue sample from brain; ti.o: sample from non-brain tissue; cs.p: cell-type-specific progenitor sample; \\
9. region: Anatomical Region (large structure); \\
10. transmitter: Transmitter; \\
11. allenregion: Region using Allen Reference Atlas notation; \\
12. num\_cells: Number of cells used in the sample; \\
13. age\_(day): Postnatal age (in days) of the mouse; \\
14. sex: Sex of the mouse; \\
15. weight\_(g): Weight (g) of the mouse; \\
16. ercc(10\^-5 dilution ul): Amount of added ERCC in ul. ($10^{-5}$ diluted); \\
17. ercc\_mix: Which ERCC mix is used; \\
18. adaptor: Which Illumina (Solexa) sequencing adaptor is used; \\
19. total\_reads: Total number of sequencing reads; \\
20. total\_wo\_ERCC: Total number of sequencing reads without reads mapping to ERCC; \\
21. read\_length: Sequencing read length; \\
22. ercc\%: Percentage of ERCC reads; \\
23. ribosomal\_etc\%: Percentage of reads mapping to ribosomal or other abundant sequences (phiX, polyC, polyA); \\
24. unmapped\_reads\%: Percentage of reads not mapped to mm10 genome; \\
25. unique\_reads\%: Percentage of reads uniquely mapped; \\
26. nonunique\_reads\%: Percentage of non-uniquely mapped reads; \\
27. short\_insert\%: Percentage of short (<30bp) reads; \\
28. mapped\_reads: Number of mapped reads; \\
29. comments: Comments; \\

\textbf{Supplementary File 3} Table listing public tissue samples used in analyses. 