\textbf{Mechanisms contributing to low noise and high information content of Homeobox TFs.} \textbf{(A)} Example expression patterns of a LIM class homeobox TF (Lhx1) and a calcium binding protein (Calb2) with similar overall expression levels. Sample key as in Figure 1 Supplements 1-3. \textbf{(B)} (upper) OFF state noise (defined as standard deviation (std) of samples with FPKM<1) plotted against maximum expression. (lower) HUGO gene groups enriched in the region indicated by red dashed lines in the upper panel (see Figure 5 Supplement 1 for PANTHER and Gene Ontology enrichments). \textbf{(C)} Average (replicate n=2) ATAC-seq profiles for the genes shown in A. Some peaks are truncated. Expression levels are plotted at right (grey bars). \textbf{(D)} Length-normalized ATAC profile for genes with high ($>$ 0.3, blue dashed box in B, n=853) and low ($<$ 0.2, red dashed box in B, n=1643) off state expression noise. \textbf{(E)} Each circle represents the separability of cell population calculated using each of PANTHER families. Separability is a measure of distance between cell populations (defined as the average of 1- Pearson’s corr. coef.) calculated using a set of genes, indicating on average how well a set of genes can distinguish (separate) cell populations. Since dispersion of separability decreases with family size, results are compared to separability calculated from randomly sampled groups of genes (green solid lines: mean and std; green dashed lines: 99\% confidence interval). Z-scores: homeobox TF: 17.4, GPCR: 16.1, receptor: 13.1 and signaling molecule: 11.2.