%%width=0.8\columnwidth
\textbf{Long genes have a greater capacity for differential expression.}
\textbf{(A)} Black dots: DI of each gene is plotted against sorted gene length. Red dots: binned average of DIs (1000 genes per bin, sorted by length). 
\textbf{(B)} Fraction of the longest 500 genes expressed within each brain region profiled for neuronal (red bars) and non-neuronal cell types (blue bars).
\textbf{(C)} ATAC-seq peaks for \textit{Gabra5} showing different patterns of peaks for each of 7 cell types. Scale (top right) in reads per million.  Expression levels for each cell type are shown at right (gray bars).
\textbf{(D)} Black dots: number of distinct peak patterns observed across 7 ATAC-seq profiled cell types plotted against the gene length for each gene; 7 corresponds to a distinct pattern for each profiled cell type. Red dots: binned averages of black dots as in panel A. Background histograms show numbers of genes in each length bin. 
\textbf{(E)} Violin plot showing the relationship between DI and the number of different patterns of ATAC-seq peaks. Corr.coef. (0.31) is greater than that between DI and gene length (0.19; panel A).
\textbf{(F)} Average metrics for long ($\geq$100kbp) and short ($<$100kbp) neuronal genes (reproducibly expressed in neuronal cell types). 
\textbf{(G)} Separability of cell types calculated as in Figure 4E, but using long neuronal genes and short neuronal genes rather than functionally defined gene families. Z-score is 33.2 for long and 22.1 for short neuronal genes. Both are highly different from randomly sampled genes (green solid lines mean and Std. dev.; dashed lines = 99\% confidence interval), but long genes provide greater separation.
